---
title: "Sensitivity of metrics to number of resources per aggregation unit"
date: "`r format(Sys.time(), '%B, %Y')`"
fontsize: 20pt
output:
  html_document:
    toc: false
    toc_depth: '3'
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

rm(list = ls())

# load libraries
library(here)
library(tidyverse)
library(ggplot2)
library(sf)
sf_use_s2(FALSE)
library(colorspace)
library(ggpubr)

# eez-land merge shapefile
regions <- st_read("data/EEZ_land_union_v3_202003/EEZ_Land_v3_202030.shp") %>% 
  filter(POL_TYPE != "Joint regime (EEZ)",
         is.na(SOVEREIGN2),
         SOVEREIGN1 != "Republic of Mauritius",
         UNION != "Antarctica") %>% 
  dplyr::select(UNION, SOVEREIGN1)
```

Supplementary analysis comparing the metrics of interest in the paper against the number of resources present in each aggregation unit and for both aggregation levels (countries and regional trade agreements).

## Number of resources per country

Computing the number of resources modeled by country across the three sectors considered. Resources can vary from 1 to 6.

```{r nres per country, echo=FALSE, warning=FALSE, message=FALSE}

# load time-series data
pct_diff_ts_r_cs_ag <- read_csv("data/data_processing/countries_time_series_1985-2015_2070-2099_agriculture.csv")
pct_diff_ts_r_cs_fi <- read_csv("data/data_processing/countries_time-series_1985-2015_2070-2099_marine-fishery_global.csv")
pct_diff_ts_r_cs_wa <- read_csv("data/data_processing/countries_time_series_1985-2015_2070-2099_water_global.csv")
pct_diff_ts_r_cs <- rbind(pct_diff_ts_r_cs_ag, pct_diff_ts_r_cs_fi, pct_diff_ts_r_cs_wa)

nres_r <- data.frame(pct_diff_ts_r_cs) %>% 
  filter(!is.na(percent_diff),
         output_variable %in% c("rice", "wheat", "soybean", "maize", "tcblog10", "qtot")) %>% 
  group_by(regions) %>% 
  summarize(n_res = length(unique(output_variable)))

nres_r_map <- left_join(regions, nres_r, by = c("SOVEREIGN1" = "regions")) 

nres_r_map %>% 
  ggplot() + geom_sf(aes(fill = n_res)) +
  theme_bw() +
  scale_fill_continuous_sequential(palette = "Grays", limits = c(1,6)) +
  theme(legend.position = "bottom")
```

How are the metrics of interest influences by the number of resources per country?

```{r sensi country, echo=FALSE, warning=FALSE, message=FALSE}

shock_probas_r_cs_mec <- read_csv("data/short-term_change/countries_shocks_1985-2015_probas_spans_mechanisms.csv") %>% 
  filter(spans == "0.75") %>% 
  dplyr::select(-spans, -at_least_three_shocks_up, -at_least_two_shocks_up)
change_probas_r_cs_mec <- read_csv("data/long-term_change/countries_long-term_change_1985-2015_probas_mechanisms.csv") %>% 
  dplyr::select(-spatial_scale)

# global
shock_probas_g_cs_mec <- read_csv("data/short-term_change/global_shocks_1985-2015_probas_spans_mechanisms.csv") %>% 
  mutate(regions = "global") %>% 
  filter(spans == "0.75") %>% 
  dplyr::select(-spans) %>% 
  dplyr::select(names(shock_probas_r_cs_mec))
change_probas_g_cs_mec <- read_csv("data/long-term_change/global_long-term_change_1985-2015_probas_mechanisms.csv") %>% 
  mutate(regions = "global") %>% 
  dplyr::select(names(change_probas_r_cs_mec))

# rbind
shock_probas <- rbind(shock_probas_g_cs_mec, shock_probas_r_cs_mec)
long_term <- rbind(change_probas_g_cs_mec, change_probas_r_cs_mec) %>% 
  pivot_wider(names_from = "type", values_from = "probas")
dat_mec <- left_join(long_term, shock_probas, by = c("climates", "regions", "time_window")) %>% 
  filter(regions != "Antarctica")

# countries
shock_probas_r_cs <- read_csv("data/short-term_change/countries_shocks_1985-2015_probas_spans.csv") %>% 
  filter(spans == "shock_0.75") %>% 
  dplyr::select(-spans)
change_probas_r_cs <- read_csv("data/long-term_change/countries_long-term_change_1985-2015_probas.csv") %>% 
  dplyr::select(-spatial_scale)

# global
shock_probas_g_cs <- read_csv("data/short-term_change/global_shocks_1985-2015_probas_spans.csv") %>% 
  mutate(regions = "global") %>% 
  filter(spans == "shock_0.75") %>% 
  dplyr::select(-spans) %>% 
  dplyr::select(names(shock_probas_r_cs))
change_probas_g_cs <- read_csv("data/long-term_change/global_long-term_change_1985-2015_probas.csv") %>% 
  mutate(regions = "global") %>% 
  dplyr::select(names(change_probas_r_cs))

# rbind
shock_probas <- rbind(shock_probas_g_cs, shock_probas_r_cs)
long_term <- rbind(change_probas_g_cs, change_probas_r_cs) %>% 
  pivot_wider(names_from = "type", values_from = "probas")
dat <- left_join(long_term, shock_probas, by = c("climates", "regions", "time_window")) %>% 
  filter(regions != "Antarctica")

dat_mec <- left_join(dat_mec, dat, by = c("regions","climates","time_window"))

# ssp5.85
dat_mec_ssp585 <- dat_mec %>% 
  filter(climates %in% c("gfdl-esm4 ssp585","ipsl-cm6a-lr ssp585"),
         time_window==30,
         regions != "global") %>% 
  group_by(regions) %>% 
  summarize(at_least_one_25_up_down = mean(at_least_one_25_up_down, na.rm=T),
            at_least_two_25_down = mean(at_least_two_25_down, na.rm=T),
            at_least_one_shock_up_down = mean(at_least_one_shock_up_down, na.rm=T),
            at_least_two_shocks_down = mean(at_least_two_shocks_down, na.rm=T),
            at_least_two_change_25 = mean(at_least_two_change_25, na.rm = T),
            at_least_two_shocks = mean(at_least_two_shocks, na.rm = T))

nres_r_metrics <- left_join(nres_r, dat_mec_ssp585) %>% 
    rename(`Gradual compensation` = at_least_one_25_up_down,
         `Gradual synchrony` = at_least_two_25_down,
         `Abrupt compensation` = at_least_one_shock_up_down,
         `Abrupt synchrony` = at_least_two_shocks_down,
         `Gradual change` = at_least_two_change_25,
         `Abrupt change` = at_least_two_shocks) %>% 
    pivot_longer(3:8, names_to = "metric", values_to = "value")

ggplot(nres_r_metrics) +
  geom_boxplot(aes(x = as.factor(n_res), y = value)) +
  facet_wrap(~ factor(metric, 
                      levels = c("Gradual change","Abrupt change",
                                 "Gradual compensation","Abrupt compensation",
                                 "Gradual synchrony","Abrupt synchrony")),
             nrow = 3) +
  theme_bw() +
  xlab("Number of resources")
```

## Number of resources per regional trade agreement

```{r nres per rta, echo=FALSE, warning=FALSE, message=FALSE}

                # load time-series data
                pct_diff_ts_rta_cs_ag <- read_csv("data/data_processing/rta_time-series_1985-2015_2070-2099_agriculture.csv") %>% 
                  dplyr::rename(regions = rtas)
                pct_diff_ts_rta_cs_fi <- read_csv("data/data_processing/rta_time-series_1985-2015_2070-2099_marine-fishery_global.csv")
                pct_diff_ts_rta_cs_wa <- read_csv("data/data_processing/rta_time-series_1985-2015_2070-2099_water_global.csv")
                pct_diff_ts_rta_cs <- rbind(pct_diff_ts_rta_cs_ag, pct_diff_ts_rta_cs_fi, pct_diff_ts_rta_cs_wa)

                nres_rta <- pct_diff_ts_rta_cs %>% 
                  filter(!is.na(percent_diff),output_variable %in% c("rice", "wheat", "soybean", "maize", "tcblog10", "qtot")) %>% 
                  group_by(regions) %>% 
                  summarize(n_res = length(unique(output_variable)))

                #### A. load data for short- and long-term changes ----
                # countries
                shock_probas_r_cs <- read_csv("data/short-term_change/countries_shocks_1985-2015_probas_spans.csv") %>% 
                  mutate(spatial_scale = "regions") %>% 
                  filter(spans == "shock_0.75") %>% 
                  dplyr::select(-spans)
                change_probas_r_cs <- read_csv("data/long-term_change/countries_long-term_change_1985-2015_probas.csv") %>% 
                  filter(time_window == 30)

                # global
                shock_probas_g_cs <- read_csv("data/short-term_change/global_shocks_1985-2015_probas_spans.csv") %>% 
                  mutate(regions = "global",
                         spatial_scale = "global") %>% 
                  filter(spans == "shock_0.75") %>% 
                  dplyr::select(names(shock_probas_r_cs))
                change_probas_g_cs <- read_csv("data/long-term_change/global_long-term_change_1985-2015_probas.csv") %>% 
                  mutate(regions = "global",
                         spatial_scale = "global") %>% 
                  filter(time_window == 30) %>% 
                  dplyr::select(names(change_probas_r_cs))

                # rtas
                shock_probas_rta_cs <- read_csv("data/short-term_change/rta_shocks_1985-2015_probas_spans.csv") %>% 
                  mutate(spatial_scale = "rta") %>% 
                  filter(spans == "shock_0.75") %>% 
                  dplyr::select(names(shock_probas_r_cs))
                change_probas_rta_cs <- read_csv("data/long-term_change/rta_long-term_change_1985-2015_probas.csv") %>% 
                  filter(time_window == 30) %>% 
                  dplyr::select(names(change_probas_r_cs))

                # rbind
                shock_probas <- rbind(shock_probas_g_cs, shock_probas_r_cs, shock_probas_rta_cs)
                long_term <- rbind(change_probas_g_cs, change_probas_r_cs, change_probas_rta_cs) %>% 
                  pivot_wider(names_from = "type", values_from = "probas")
                dat <- left_join(long_term, shock_probas, by = c("climates", "regions", "spatial_scale")) %>% 
                  filter(regions != "Antarctica")

                #### B. load data for compensatory and synchronous mechanisms ----
                # countries
                shock_probas_r_cs_mec <- read_csv("data/short-term_change/countries_shocks_1985-2015_probas_spans_mechanisms.csv") %>% 
                  mutate(spatial_scale = "regions") %>% 
                  filter(spans == "0.75") %>% 
                  dplyr::select(-spans, -at_least_two_shocks_up, -at_least_three_shocks_up)
                change_probas_r_cs_mec <- read_csv("data/long-term_change/countries_long-term_change_1985-2015_probas_mechanisms.csv") %>% 
                  filter(time_window == 30) %>% 
                  dplyr::select(-time_window)

                # global
                shock_probas_g_cs_mec <- read_csv("data/short-term_change/global_shocks_1985-2015_probas_spans_mechanisms.csv") %>% 
                  mutate(regions = "global",
                         spatial_scale = "global") %>% 
                  filter(spans == "0.75") %>% 
                  dplyr::select(names(shock_probas_r_cs_mec))
                change_probas_g_cs_mec <- read_csv("data/long-term_change/global_long-term_change_1985-2015_probas_mechanisms.csv") %>% 
                  mutate(regions = "global",
                         saptial_scale = "global") %>% 
                  filter(time_window == 30) %>% 
                  dplyr::select(names(change_probas_r_cs_mec))

                # rta
                shock_probas_rta_cs_mec <- read_csv("data/short-term_change/rta_shocks_1985-2015_probas_spans_mechanisms.csv") %>% 
                  mutate(spatial_scale = "rta") %>% 
                  filter(spans == "0.75") %>% 
                  dplyr::select(names(shock_probas_r_cs_mec))
                change_probas_rta_cs_mec <- read_csv("data/long-term_change/rta_long-term_change_1985-2015_probas_mechanisms.csv")  %>% 
                  filter(time_window == 30) %>% 
                  dplyr::select(names(change_probas_r_cs_mec))

                # rbind
                shock_probas <- rbind(shock_probas_g_cs_mec, shock_probas_r_cs_mec, shock_probas_rta_cs_mec)
                long_term <- rbind(change_probas_g_cs_mec, change_probas_r_cs_mec, change_probas_rta_cs_mec) %>% 
                  pivot_wider(names_from = "type", values_from = "probas")
                dat_mec <- left_join(long_term, shock_probas, by = c("climates", "regions", "spatial_scale")) %>% 
                  filter(regions != "Antarctica")


                #### C. Data across metrics ----
                # cbind both datasets
                dat_ecology_rta <- left_join(dat, dat_mec, by = c("regions", "climates", "spatial_scale")) %>% 
                  dplyr::select(regions, climates, spatial_scale,
                                at_least_two_change_25, at_least_two_shocks,
                                at_least_two_25_down, at_least_one_25_up_down,
                                at_least_two_shocks_down, at_least_one_shock_up_down) %>%
                  mutate(at_least_one_25_up_down = 1-at_least_one_25_up_down,
                         at_least_one_shock_up_down = 1-at_least_one_shock_up_down) %>%
                  pivot_longer(4:9, names_to = "type", values_to = "proba") %>% 
                  mutate(scale = ifelse(str_detect(type, "shock")==TRUE, "abrupt", "gradual"),
                         mechanism = ifelse(str_detect(type, "down")==TRUE, "synchrony", "stability"),
                         mechanism = ifelse(str_detect(type, "up_down")==TRUE, "compensation", mechanism)) %>% 
                  filter(spatial_scale == "rta",
                         climates %in% c("gfdl-esm4 ssp585","ipsl-cm6a-lr ssp585")) %>% 
                  group_by(regions, spatial_scale, type, scale, mechanism) %>% 
                  summarize(proba = mean(proba, na.rm=T))
                dat_ecology_rta$regions <- as.numeric(as.vector(dat_ecology_rta$regions))

nres_rta_metrics <- left_join(nres_rta, dat_ecology_rta) %>% 
  mutate(mechanism = ifelse(mechanism == "stability","change",mechanism),
         type = paste(scale, mechanism, sep = " "))

ggplot(nres_rta_metrics) +
  geom_boxplot(aes(x = as.factor(n_res), y = proba)) +
  facet_wrap(~ factor(type,
             levels = c("gradual change","abrupt change",
                        "gradual compensation","abrupt compensation",
                        "gradual synchrony","abrupt synchrony")),
             nrow = 3) +
  theme_bw() +
  xlab("Number of resources")
```

## Make summary figure

```{r summary figure, fig.width=10, fig.height=12}

map <- nres_r_map %>% 
  ggplot() + geom_sf(aes(fill = n_res)) +
  theme_bw() +
  scale_fill_continuous_sequential(palette = "Grays", limits = c(1,6))

box_cnt <- ggplot(nres_r_metrics) +
  geom_boxplot(aes(x = as.factor(n_res), y = value)) +
  facet_wrap(~ factor(metric, 
                      levels = c("Gradual change","Abrupt change",
                                 "Gradual compensation","Abrupt compensation",
                                 "Gradual synchrony","Abrupt synchrony")),
             nrow = 3) +
  theme_bw() +
  xlab("Number of resources") +
  ylab("Probability")

box_rta <- ggplot(nres_rta_metrics) +
  geom_boxplot(aes(x = as.factor(n_res), y = proba)) +
  facet_wrap(~ factor(type,
             levels = c("gradual change","abrupt change",
                        "gradual compensation","abrupt compensation",
                        "gradual synchrony","abrupt synchrony")),
             nrow = 3) +
  theme_bw() +
  xlab("Number of resources") +
  ylab("Probability")

ggarrange(map, 
          ggarrange(box_cnt, box_rta, ncol = 2,
                    labels = c("CNT","RTA")),
          nrow = 2, labels = "Number of resources",
          heights = c(1,2), widths = c(1,1))
```

## 
