---
title: "cross sector shocks"
date: "`r format(Sys.time(), '%B, %Y')`"
fontsize: 20pt
output:
  html_document:
    toc: false
    toc_depth: '3'
    df_print: paged
---

```{r setup, include=FALSE, warnings = FALSE, message = FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
# load libraries
library(here)
library(tidyverse)
library(ggplot2)

# load data
shock_ts_g_cs <- read_csv("/Users/auroremaureaud/Library/CloudStorage/Dropbox/Minerva 2021 Climate Politics/Aurore/Resource synthesis/future_resource_change/data/short-term_change/global_shocks_time_series_2010-2019.csv")
```

## Shock probability across models

To get the yearly shock probability for each combination of sector, climate model, climate experiment, and output_variable, we:

-   applied a smoother (loess R function) per ecological model

-   extract the z-scored residuals

-   counted the number of shocks (beyond 2SD of the mean) across the time-series, and whether the shocks are from increases or decreases in the resource level

-   calculated the mean number of shocks across ecological models, as a way to aggregate the model ensembles

Below is the example of maize globally across climates:

```{r maizeshocks, fig.width=10, warning=FALSE, message=FALSE}
shock_ts_g_cs %>% 
  filter(output_variable != "ptotww",
         output_variable == "maize") %>%
  mutate(shock = ifelse(is.na(shock),0,shock),
         shock_up = ifelse(is.na(shock_up),0,shock_up),
         shock_down = ifelse(is.na(shock_down),0,shock_down)) %>% 
  mutate(climates = paste(climate_model, experiment_climate)) %>% 
  group_by(climates, years, output_variable) %>% 
  summarize(shock = mean(shock),
            shock_down = mean(shock_down),
            shock_up = mean(shock_up)) %>% 
  pivot_longer(c("shock","shock_down","shock_up"), names_to = "shock_type",values_to = "p_shock") %>% 
  ggplot(aes(x = years, y = p_shock, fill = shock_type)) + 
  geom_bar(stat = "identity", position = "dodge") +
  theme_bw() +
  ylim(0,1) + ylab("Shock probability") +
  scale_fill_manual(breaks = c("shock","shock_down","shock_up"), 
                    labels = c("shock","down","up"),
                    values = c("grey50","mediumaquamarine","lightsalmon")) +
  facet_wrap(~ climates)
```

## Shock probability for multiple resources

Shock probabilities for GFDL-ESM4 and SSP585 across sectors and modelled resources:

```{r shockprobabilities, fig.width=10, echo=FALSE, warning=FALSE, message=FALSE}

shock_ts_g_cs %>% 
  filter(output_variable != "ptotww") %>%
  mutate(shock = ifelse(is.na(shock),0,shock),
         shock_up = ifelse(is.na(shock_up),0,shock_up),
         shock_down = ifelse(is.na(shock_down),0,shock_down)) %>% 
  mutate(climates = paste(climate_model, experiment_climate)) %>% 
  filter(climates == "gfdl-esm4 ssp585") %>% 
  group_by(climates, years, output_variable) %>% 
  summarize(shock = mean(shock),
            shock_down = mean(shock_down),
            shock_up = mean(shock_up)) %>% 
  pivot_longer(c("shock","shock_down","shock_up"), names_to = "shock_type",values_to = "p_shock") %>% 
  ggplot(aes(x = years, y = p_shock, fill = shock_type)) + 
  geom_bar(stat = "identity", position = "dodge") +
  theme_bw() +
  ylab("Shock probability") +
  scale_fill_manual(breaks = c("shock","shock_down","shock_up"), 
                    labels = c("shock","down","up"),
                    values = c("grey50","mediumaquamarine","lightsalmon")) +
  facet_wrap(~ output_variable)
```

## Cross-sector shock probability

To get the cross-sector shock occurrence probability, we calculate the shock probability from multiple sector together. For this, we selected maize, soybean, rice, wheat, tcblog10, and qtot, and we compute the cross-sector shock probability as follows:

-   probability that there is at least one shock per year: P(X\>=1) = 1-P(X=0)

-   probability that there are at least two shocks per year: P(X\>=2) = P(X\>=1)-P(X=1)

-   probability that there are at least three shocks per year: P(X\>=3) = P(X\>=2)-P(X=3)

The plot below shows the cross-sector probability across climates globally:

```{r crosssectorprobability, fig.width=10, warning=FALSE, message=FALSE}

shock_ts_g_cs %>% 
  filter(!output_variable %in% c("tws","tcb","ptotww")) %>%
  mutate(shock = ifelse(is.na(shock),0,shock),
         shock_up = ifelse(is.na(shock_up),0,shock_up),
         shock_down = ifelse(is.na(shock_down),0,shock_down)) %>% 
  mutate(climates = paste(climate_model, experiment_climate)) %>% 
  group_by(climates, years, output_variable) %>% 
  summarize(shock = mean(shock),
            shock_down = mean(shock_down),
            shock_up = mean(shock_up)) %>% 
  dplyr::select(climates, years, output_variable, shock) %>% 
  pivot_wider(names_from = output_variable, values_from = shock) %>% 
  mutate(
         # probability there is at least one shock per year
         at_least_one_shock = 1-(1-maize)*(1-qtot)*(1-rice)*(1-soybean)*(1-tcblog10)*(1-wheat),
         # probability there is one shock per year
         one_shock = maize*(1-qtot)*(1-rice)*(1-soybean)*(1-tcblog10)*(1-wheat) + 
                     qtot*(1-maize)*(1-rice)*(1-tcblog10)*(1-soybean)*(1-wheat) +
                     rice*(1-qtot)*(1-maize)*(1-tcblog10)*(1-soybean)*(1-wheat) +
                     soybean*(1-maize)*(1-qtot)*(1-rice)*(1-tcblog10)*(1-wheat) +
                     tcblog10*(1-maize)*(1-qtot)*(1-rice)*(1-soybean)*(1-wheat) +
                     wheat*(1-maize)*(1-qtot)*(1-rice)*(1-tcblog10)*(1-soybean),
         # probability there are at least two shocks per year
         at_least_two_shocks = at_least_one_shock-one_shock,
         # probability there are two shocks per year
         two_shocks = maize*qtot*(1-rice)*(1-soybean)*(1-tcblog10)*(1-wheat) +
                      maize*rice*(1-qtot)*(1-soybean)*(1-tcblog10)*(1-wheat) +
                      maize*soybean*(1-qtot)*(1-rice)*(1-tcblog10)*(1-wheat) +
                      maize*tcblog10*(1-qtot)*(1-rice)*(1-soybean)*(1-wheat) +
                      maize*wheat*(1-qtot)*(1-rice)*(1-soybean)*(1-tcblog10) +
                      qtot*rice*(1-maize)*(1-soybean)*(1-tcblog10)*(1-wheat) +
                      qtot*soybean*(1-maize)*(1-rice)*(1-tcblog10)*(1-wheat) +
                      qtot*tcblog10*(1-maize)*(1-rice)*(1-soybean)*(1-wheat) +
                      qtot*wheat*(1-maize)*(1-rice)*(1-tcblog10)*(1-soybean) +
                      rice*soybean*(1-maize)*(1-qtot)*(1-tcblog10)*(1-wheat) +
                      rice*tcblog10*(1-maize)*(1-soybean)*(1-qtot)*(1-wheat) +
                      rice*wheat*(1-maize)*(1-soybean)*(1-qtot)*(1-tcblog10) +
                      soybean*tcblog10*(1-maize)*(1-qtot)*(1-rice)*(1-wheat) +
                      soybean*wheat*(1-maize)*(1-qtot)*(1-rice)*(1-tcblog10) +
                      tcblog10*wheat*(1-maize)*(1-soybean)*(1-rice)*(1-qtot),
         # probability there are at least three shocks per year
         at_least_three_shocks = at_least_two_shocks-two_shocks) %>% 
  dplyr::select(climates, years, at_least_one_shock, at_least_two_shocks, at_least_three_shocks) %>% 
  pivot_longer(c("at_least_one_shock","at_least_two_shocks","at_least_three_shocks"), names_to = "shock_type",values_to = "p_shock") %>% 
  ggplot(aes(x = years, y = p_shock, fill = shock_type)) +
  #scale_y_log10() +
  geom_bar(stat = "identity", position = "dodge") +
  theme_bw() +
  ylab("Cross-sector shock probability") +
  scale_fill_manual(breaks = c("at_least_one_shock","at_least_two_shocks","at_least_three_shocks"), 
                    labels = c("at least 1","at least 2","at least 3"),
                    values = c("gold","darkorange","firebrick3")) +
  facet_wrap(~ climates)
```

Summary of cross-sector shock occurrence probability over the last decade (2080-2100) across climates and cross-sector shock types:

```{r crosssectorstats, echo=FALSE, warning=FALSE, message=FALSE, fig.height=4}

shock_ts_g_cs %>% 
  filter(!output_variable %in% c("tws","tcb","ptotww")) %>%
  mutate(shock = ifelse(is.na(shock),0,shock),
         shock_up = ifelse(is.na(shock_up),0,shock_up),
         shock_down = ifelse(is.na(shock_down),0,shock_down)) %>% 
  mutate(climates = paste(climate_model, experiment_climate)) %>% 
  group_by(climates, years, output_variable) %>% 
  summarize(shock = mean(shock),
            shock_down = mean(shock_down),
            shock_up = mean(shock_up)) %>% 
  dplyr::select(climates, years, output_variable, shock) %>% 
  pivot_wider(names_from = output_variable, values_from = shock) %>% 
  mutate(
         # probability there is at least one shock per year
         at_least_one_shock = 1-(1-maize)*(1-qtot)*(1-rice)*(1-soybean)*(1-tcblog10)*(1-wheat),
         # probability there is one shock per year
         one_shock = maize*(1-qtot)*(1-rice)*(1-soybean)*(1-tcblog10)*(1-wheat) + 
                     qtot*(1-maize)*(1-rice)*(1-tcblog10)*(1-soybean)*(1-wheat) +
                     rice*(1-qtot)*(1-maize)*(1-tcblog10)*(1-soybean)*(1-wheat) +
                     soybean*(1-maize)*(1-qtot)*(1-rice)*(1-tcblog10)*(1-wheat) +
                     tcblog10*(1-maize)*(1-qtot)*(1-rice)*(1-soybean)*(1-wheat) +
                     wheat*(1-maize)*(1-qtot)*(1-rice)*(1-tcblog10)*(1-soybean),
         # probability there are at least two shocks per year
         at_least_two_shocks = at_least_one_shock-one_shock,
         # probability there are two shocks per year
         two_shocks = maize*qtot*(1-rice)*(1-soybean)*(1-tcblog10)*(1-wheat) +
                      maize*rice*(1-qtot)*(1-soybean)*(1-tcblog10)*(1-wheat) +
                      maize*soybean*(1-qtot)*(1-rice)*(1-tcblog10)*(1-wheat) +
                      maize*tcblog10*(1-qtot)*(1-rice)*(1-soybean)*(1-wheat) +
                      maize*wheat*(1-qtot)*(1-rice)*(1-soybean)*(1-tcblog10) +
                      qtot*rice*(1-maize)*(1-soybean)*(1-tcblog10)*(1-wheat) +
                      qtot*soybean*(1-maize)*(1-rice)*(1-tcblog10)*(1-wheat) +
                      qtot*tcblog10*(1-maize)*(1-rice)*(1-soybean)*(1-wheat) +
                      qtot*wheat*(1-maize)*(1-rice)*(1-tcblog10)*(1-soybean) +
                      rice*soybean*(1-maize)*(1-qtot)*(1-tcblog10)*(1-wheat) +
                      rice*tcblog10*(1-maize)*(1-soybean)*(1-qtot)*(1-wheat) +
                      rice*wheat*(1-maize)*(1-soybean)*(1-qtot)*(1-tcblog10) +
                      soybean*tcblog10*(1-maize)*(1-qtot)*(1-rice)*(1-wheat) +
                      soybean*wheat*(1-maize)*(1-qtot)*(1-rice)*(1-tcblog10) +
                      tcblog10*wheat*(1-maize)*(1-soybean)*(1-rice)*(1-qtot),
         # probability there are at least three shocks per year
         at_least_three_shocks = at_least_two_shocks-two_shocks) %>% 
  dplyr::select(climates, years, at_least_one_shock, at_least_two_shocks, at_least_three_shocks) %>% 
  pivot_longer(c("at_least_one_shock","at_least_two_shocks","at_least_three_shocks"), names_to = "shock_type",values_to = "p_shock") %>% 
  mutate(p_shock = ifelse(p_shock<0, 0, p_shock)) %>% 
  filter(years>2079) %>% 
  data.frame() %>% 
  ggplot(aes(y = p_shock, x = climates)) + 
  geom_boxplot() +
  theme_bw() + theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab("Cross-sector shock probability") +
  scale_fill_manual(breaks = c("at_least_one_shock","at_least_two_shocks","at_least_three_shocks"), 
                    labels = c("1+","2+","3+"),
                    values = c("gold","darkorange","firebrick3")) +
  facet_wrap(~ shock_type)
```

## Cross-sector shock probability across years

To get the cross-sector occurrence probability across years, we calculate the cross-sector probability over the last 30 years that there is:

-   at least one year with at least one shock in the same year: P(X\>=1) = 1-P(X=0) = 1-(1-Px\>=1,y1)\**...\**(1-Px\>=1,y30)

-   at least one year with at least two shocks in the same year: P(X\>=2) = P(X\>=1)-P(X=1) = P(X\>=1) - (Px\>=1,y1\*(1-Px\>=1,y2)\*...\*(1-Px\>=1,y30) +...+ Px\>=1,y30\*(1-Px\>=1,y1)\*...\*(1-Px\>=1,y29)

-   at least one year with at least three shocks in the same year:

```{r crosssectorcrosstimeprobability, fig.width=10, warning=FALSE, message=FALSE}

xx <- shock_ts_g_cs %>% 
  filter(!output_variable %in% c("tws","tcb","ptotww")) %>%
  mutate(shock = ifelse(is.na(shock),0,shock),
         shock_up = ifelse(is.na(shock_up),0,shock_up),
         shock_down = ifelse(is.na(shock_down),0,shock_down)) %>% 
  mutate(climates = paste(climate_model, experiment_climate)) %>% 
  group_by(climates, years, output_variable) %>% 
  summarize(shock = mean(shock),
            shock_down = mean(shock_down),
            shock_up = mean(shock_up)) %>% 
  dplyr::select(climates, years, output_variable, shock) %>% 
  pivot_wider(names_from = output_variable, values_from = shock) %>% 
  mutate(
         # probability there is at least one shock per year
         at_least_one_shock = 1-(1-maize)*(1-qtot)*(1-rice)*(1-soybean)*(1-tcblog10)*(1-wheat),
         # probability there is one shock per year
         one_shock = maize*(1-qtot)*(1-rice)*(1-soybean)*(1-tcblog10)*(1-wheat) + 
                     qtot*(1-maize)*(1-rice)*(1-tcblog10)*(1-soybean)*(1-wheat) +
                     rice*(1-qtot)*(1-maize)*(1-tcblog10)*(1-soybean)*(1-wheat) +
                     soybean*(1-maize)*(1-qtot)*(1-rice)*(1-tcblog10)*(1-wheat) +
                     tcblog10*(1-maize)*(1-qtot)*(1-rice)*(1-soybean)*(1-wheat) +
                     wheat*(1-maize)*(1-qtot)*(1-rice)*(1-tcblog10)*(1-soybean),
         # probability there are at least two shocks per year
         at_least_two_shocks = at_least_one_shock-one_shock,
         # probability there are two shocks per year
         two_shocks = maize*qtot*(1-rice)*(1-soybean)*(1-tcblog10)*(1-wheat) +
                      maize*rice*(1-qtot)*(1-soybean)*(1-tcblog10)*(1-wheat) +
                      maize*soybean*(1-qtot)*(1-rice)*(1-tcblog10)*(1-wheat) +
                      maize*tcblog10*(1-qtot)*(1-rice)*(1-soybean)*(1-wheat) +
                      maize*wheat*(1-qtot)*(1-rice)*(1-soybean)*(1-tcblog10) +
                      qtot*rice*(1-maize)*(1-soybean)*(1-tcblog10)*(1-wheat) +
                      qtot*soybean*(1-maize)*(1-rice)*(1-tcblog10)*(1-wheat) +
                      qtot*tcblog10*(1-maize)*(1-rice)*(1-soybean)*(1-wheat) +
                      qtot*wheat*(1-maize)*(1-rice)*(1-tcblog10)*(1-soybean) +
                      rice*soybean*(1-maize)*(1-qtot)*(1-tcblog10)*(1-wheat) +
                      rice*tcblog10*(1-maize)*(1-soybean)*(1-qtot)*(1-wheat) +
                      rice*wheat*(1-maize)*(1-soybean)*(1-qtot)*(1-tcblog10) +
                      soybean*tcblog10*(1-maize)*(1-qtot)*(1-rice)*(1-wheat) +
                      soybean*wheat*(1-maize)*(1-qtot)*(1-rice)*(1-tcblog10) +
                      tcblog10*wheat*(1-maize)*(1-soybean)*(1-rice)*(1-qtot),
         # probability there are at least three shocks per year
         at_least_three_shocks = at_least_two_shocks-two_shocks) %>% 
  dplyr::select(climates, years, at_least_one_shock, at_least_two_shocks, at_least_three_shocks) %>% 
  filter(years>2089) %>% 
  ggplot(aes(x = years, y = p_shock, fill = shock_type)) +
  #scale_y_log10() +
  geom_bar(stat = "identity", position = "dodge") +
  theme_bw() +
  ylab("Cross-sector shock probability") +
  scale_fill_manual(breaks = c("at_least_one_shock","at_least_two_shocks","at_least_three_shocks"), 
                    labels = c("1+","2+","3+"),
                    values = c("gold","darkorange","firebrick3")) +
  facet_wrap(~ climates)
```
