---
title: "Probability Independence for the compensation mechanisms"
date: "`r format(Sys.time(), '%B, %Y')`"
fontsize: 20pt
output:
  html_document:
    toc: false
    toc_depth: '3'
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)

# load libraries
library(here)
library(tidyverse)
library(ggplot2)
library(gridExtra)
library(GGally)
library(ggpubr)
```

# Check for probability independence

The compensation probabilities rely on the product between probability of at least one resource going up and at least one resource going down. However, this formula is only correct if the two probabilities are independent, which may not be the case. Therefore, we check for independence for probabilities of at least three resources changing (up or down) and see whether they correlate.

## Shock probabilities

```{r, warning=FALSE, message=FALSE}

shock_proba_r_cs_mec <- read_csv("/Users/aurore.maureaud/Documents/GitHub/cross-sector_resource_change/data/short-term_change/countries_shocks_1985-2015_probas_spans_mechanisms.csv")

ggplot(shock_proba_r_cs_mec, aes(x = at_least_two_shocks_up, y = at_least_two_shocks_down, colour = as.factor(spans))) +
  geom_point(alpha = 0.5) +
  theme_bw() +
  facet_wrap(~climates, scales = "free") +
  geom_abline(aes(intercept = 0, slope = 1)) +
  theme(text = element_text(size = 16)) +
  ggtitle("Relationships p(2 shocks up) and p(2 shocks down)")

ggplot(shock_proba_r_cs_mec, aes(x = at_least_three_shocks_up, y = at_least_three_shocks_down, colour = as.factor(spans))) +
  geom_point(alpha = 0.5) +
  theme_bw() +
  facet_wrap(~climates, scales = "free") +
  geom_abline(aes(intercept = 0, slope = 1)) +
  theme(text = element_text(size = 16)) +
  ggtitle("Relationships p(3 shocks up) and p(3 shocks down)")

ggpairs(shock_proba_r_cs_mec, columns = c(4,10), aes(colour = climates)) +
  theme_bw() + theme(text = element_text(size = 16)) +
  ggtitle("Correlation p(2 shocks up) and p(2 shocks down)")
ggpairs(shock_proba_r_cs_mec, columns = c(5,11), aes(colour = climates)) +
  theme_bw() + theme(text = element_text(size = 16)) +
  ggtitle("Correlation p(3 shocks up) and p(3 shocks down)")
```

## Gradual change probabilities

```{r, warning=FALSE, message=FALSE}

change_proba_r_cs_mec <- read_csv("/Users/aurore.maureaud/Documents/GitHub/cross-sector_resource_change/data/long-term_change/countries_long-term_change_1985-2015_probas_mechanisms.csv") %>% 
  filter(time_window == 30,
         str_detect(type, "25")==TRUE,
         str_detect(type, "up_down")==FALSE,
         str_detect(type, "one")==FALSE) %>% 
  pivot_wider(names_from = type, values_from = probas)
  
ggplot(change_proba_r_cs_mec, aes(x = at_least_two_25_up, y = at_least_two_25_down)) +
  geom_point(alpha = 0.5) +
  theme_bw() +
  facet_wrap(~climates, scales = "free") + xlim(0,1) + ylim(0,1) +
  geom_abline(aes(intercept = 0, slope = 1)) +
  theme(text = element_text(size = 16)) +
  ggtitle("Relationships p(2 increases) and p(2 decreases)")

ggplot(change_proba_r_cs_mec, aes(x = at_least_three_25_up, y = at_least_three_25_down)) +
  geom_point(alpha = 0.5) +
  theme_bw() +
  facet_wrap(~climates, scales = "free") + xlim(0,1) + ylim(0,1) +
  geom_abline(aes(intercept = 0, slope = 1)) +
  theme(text = element_text(size = 16)) +
  ggtitle("Relationships p(3 increases) and p(3 decreases)")

ggpairs(change_proba_r_cs_mec, columns = c(5,7), aes(colour = climates)) +
  theme_bw() + theme(text = element_text(size = 16)) +
  ggtitle("Correlation p(3 increases) and p(3 decreases)")
ggpairs(change_proba_r_cs_mec, columns = c(6,8), aes(colour = climates)) +
  theme_bw() + theme(text = element_text(size = 16)) +
  ggtitle("Correlation p(3 increases) and p(3 decreases)")
```
Saving supplementary figures

```{r supplementary figure 11, warning=FALSE, message=FALSE}

two_shocks <- shock_proba_r_cs_mec %>% 
  filter(spans == "0.75") %>% 
  ggplot(aes(x = at_least_two_shocks_up, y = at_least_two_shocks_down, color = climates)) +
  geom_point(alpha = 0.5) +
  theme_bw() +
  geom_abline(aes(intercept = 0, slope = 1)) +
  theme(text = element_text(size = 12)) +
    scale_color_manual(values = c("plum","purple4","lightskyblue","blue"),
                    name = "Climate models and scenarios") +
  xlab("proba. 2 increasing shocks") +
  ylab("proba. 2 decreasing shocks")

three_shocks <- shock_proba_r_cs_mec %>% 
  filter(spans == "0.75") %>% 
  ggplot(aes(x = at_least_three_shocks_up, y = at_least_three_shocks_down, colour=climates)) +
  geom_point(alpha = 0.5) +
  theme_bw() +
  geom_abline(aes(intercept = 0, slope = 1)) +
  theme(text = element_text(size = 12)) +
    scale_color_manual(values = c("plum","purple4","lightskyblue","blue"),
                    name = "Climate models and scenarios") +
  xlab("proba. 3 increasing shocks") +
  ylab("proba. 3 decreasing shocks")

png(paste0("/Users/aurore.maureaud/Documents/GitHub/cross-sector_resource_change/figures/supplementary_figure_11.png"), width = 10*175, height = 6*200, res = 200)
ggarrange(two_shocks, three_shocks, nrow = 2,
          labels = c("A","B"))
dev.off()

```


```{r supplementary figure 12, warning=FALSE, message=FALSE}
two_res <- ggplot(change_proba_r_cs_mec, aes(x = at_least_two_25_up, y = at_least_two_25_down, color = climates)) +
  geom_point(alpha = 0.5) +
  theme_bw() +
  xlim(0,1) + ylim(0,1) +
  geom_abline(aes(intercept = 0, slope = 1)) +
  theme(text = element_text(size = 12)) +
  theme(text = element_text(size = 12)) +
    scale_color_manual(values = c("plum","purple4","lightskyblue","blue"),
                    name = "Climate models and scenarios") +
  xlab("proba. 2 increasing res.") +
  ylab("proba. 3 decreasing res.")

three_res <- ggplot(change_proba_r_cs_mec, aes(x = at_least_three_25_up, y = at_least_three_25_down, color = climates)) +
  geom_point(alpha = 0.5) +
  theme_bw() +
  xlim(0,1) + ylim(0,1) +
  geom_abline(aes(intercept = 0, slope = 1)) +
  theme(text = element_text(size = 12)) +
    theme(text = element_text(size = 12)) +
    scale_color_manual(values = c("plum","purple4","lightskyblue","blue"),
                    name = "Climate models and scenarios") +
  xlab("proba. 3 increasing res.") +
  ylab("proba. 3 decreasing res.")

png(paste0("/Users/aurore.maureaud/Documents/GitHub/cross-sector_resource_change/figures/supplementary_figure_12.png"), width = 10*175, height = 6*200, res = 200)
ggarrange(two_res, three_res, nrow = 2,
          labels = c("A","B"))
dev.off()

```