---
title: "Compare resource change metrics"
date: "`r format(Sys.time(), '%B, %Y')`"
fontsize: 20pt
output:
  html_document:
    toc: false
    toc_depth: '3'
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)

# load libraries
library(here)
library(tidyverse)
library(ggplot2)
library(gridExtra)

# load functions
source("/Users/auroremaureaud/Library/CloudStorage/Dropbox/Minerva 2021 Climate Politics/Aurore/Resource synthesis/future_resource_change/functions/get_probas.R")

# load data
pct_diff_ts_r_cs_ag <- read_csv("/Users/auroremaureaud/Library/CloudStorage/Dropbox/Minerva 2021 Climate Politics/Aurore/Resource synthesis/future_resource_change/data/data_processing/countries_time_series_1985-2015_2070-2099_agriculture.csv")
pct_diff_ts_r_cs_fi <- read_csv("/Users/auroremaureaud/Library/CloudStorage/Dropbox/Minerva 2021 Climate Politics/Aurore/Resource synthesis/future_resource_change/data/data_processing/countries_time-series_1985-2015_2070-2099_marine-fishery_global.csv")
pct_diff_ts_r_cs_wa <- read_csv("/Users/auroremaureaud/Library/CloudStorage/Dropbox/Minerva 2021 Climate Politics/Aurore/Resource synthesis/future_resource_change/data/data_processing/countries_time_series_1985-2015_2070-2099_water_global.csv")
ts <- rbind(pct_diff_ts_r_cs_ag, pct_diff_ts_r_cs_fi, pct_diff_ts_r_cs_wa) %>% 
  mutate(id = paste(sector, eco_model, climate_model, experiment_climate, output_variable, regions, sep = "/"))
rm(pct_diff_ts_r_cs_wa, pct_diff_ts_r_cs_fi, pct_diff_ts_r_cs_ag)
```

## Measuring long-term gradual change and short-term shocks

Data sample for agriculture maize crop yield for the CROVER crop model, GFDL-ESM4, SSP 585, and for the United States. We compare the three main metrics computed:

1.  End of the century (2070-2099) long-term % difference average compared to a fixed threshold of 5, 10, or 25% (code with gradual purple shades)
2.  End of century (2070-2099) long-term % difference average compared to the 10th percentile limit from the historical period (1985-2015), which is used as a specific threshold per simulation
3.  Years with shocks from 2SD of residuals from a smooth loess fit with span of 0.25, 0.5 or 0.75 (yellow, orange, and red fitted lines, respectively). The years and time-series points identified as experiencing a resource shock are colored in yellow, orange, or red.

```{r metrics, warning=FALSE, message=FALSE, fig.width=10}
ts_ex <- ts %>% 
  filter(output_variable == "maize",
         climate_model == "gfdl-esm4",
         experiment_climate == "ssp585",
         regions == "United States",
         eco_model == "crover")

# long-term change with threshold
avg <- mean(ts_ex$percent_diff[88:117], na.rm=T)

plot_ts_avg <- ggplot(ts_ex, aes(x = years, y = percent_diff)) + geom_point() + geom_line() +
  theme_bw() + xlab("Years") + ylab("% Difference") +
  annotate("rect", xmin = 1983, xmax = 2015, ymin = -Inf, ymax = Inf, alpha = .2) +
  annotate("rect", xmin = -Inf, xmax = Inf, ymin = -5, ymax = 0, alpha = .4, fill = "plum") +
    annotate("rect", xmin = -Inf, xmax = Inf, ymin = -10, ymax = -5, alpha = .4, fill = "orchid3") +
    annotate("rect", xmin = -Inf, xmax = Inf, ymin = -25, ymax = -10, alpha = .4, fill = "purple") +
annotate("rect", xmin = 1983, xmax = 2015, ymin = -Inf, ymax = Inf, alpha = .2) +
  geom_segment(aes(x = 2070, y = avg, xend = 2099, yend = avg), colour = "purple") +
  ggtitle("End of century average beyond % threshold") +
  theme(legend.position = "none") +
  geom_hline(aes(yintercept = 0), linetype = "dashed") +
  geom_segment(aes(x = 2085, xend = 2085, y = 0, yend = avg), arrow = arrow(length = unit(0.5, "cm")),colour = "purple", size = 0.5) +
   theme(plot.title = element_text(size = 10))
  

# count shocks
model_smooth <- loess(ts_ex$percent_diff ~ ts_ex$years, span = 0.75)
ts_ex$res_0.75 <- (model_smooth$residuals - mean(model_smooth$residuals))/sd(model_smooth$residuals)

model_smooth <- loess(ts_ex$percent_diff ~ ts_ex$years, span = 0.5)
ts_ex$res_0.5 <- (model_smooth$residuals - mean(model_smooth$residuals))/sd(model_smooth$residuals)

model_smooth <- loess(ts_ex$percent_diff ~ ts_ex$years, span = 0.25)
ts_ex$res_0.25 <- (model_smooth$residuals - mean(model_smooth$residuals))/sd(model_smooth$residuals)
rm(model_smooth)

ts_ex <- ts_ex %>% 
  mutate(shock_0.75 = ifelse(res_0.75>2, 1, 0),
         shock_0.75 = ifelse(res_0.75<(-2), 1, shock_0.75),
         shock_0.5 = ifelse(res_0.5>2, 1, 0),
         shock_0.5 = ifelse(res_0.5<(-2), 1, shock_0.5),
         shock_0.25 = ifelse(res_0.25>2, 1, 0),
         shock_0.25 = ifelse(res_0.25<(-2), 1, shock_0.25))

plot_shock0.75 <- ggplot(ts_ex, aes(x = years, y = percent_diff)) + geom_point() + geom_line() +
  theme_bw() + xlab("Years") + ylab("% Difference") +
  annotate("rect", xmin = 1983, xmax = 2015, ymin = -Inf, ymax = Inf, alpha = .2) +
  geom_point(data = ts_ex[ts_ex$shock_0.75==1,], col = "red") +
  ggtitle("Smooth with span of 0.75") +
  geom_hline(aes(yintercept = 0), linetype = "dashed")+
   theme(plot.title = element_text(size = 10)) +
  geom_smooth(span = 0.75, col = "red", level = 0.95)

plot_shock0.5 <- ggplot(ts_ex, aes(x = years, y = percent_diff)) + geom_point() + geom_line() +
  theme_bw() + xlab("Years") + ylab("% Difference") +
  annotate("rect", xmin = 1983, xmax = 2015, ymin = -Inf, ymax = Inf, alpha = .2) +
  geom_point(data = ts_ex[ts_ex$shock_0.5==1,], col = "orange") +
  ggtitle("Smooth with span of 0.5") +
  geom_hline(aes(yintercept = 0), linetype = "dashed")+
   theme(plot.title = element_text(size = 10)) +
  geom_smooth(span = 0.5, col = "orange", level = 0.95)

plot_shock0.25 <- ggplot(ts_ex, aes(x = years, y = percent_diff)) + geom_point() + geom_line() +
  theme_bw() + xlab("Years") + ylab("% Difference") +
  annotate("rect", xmin = 1983, xmax = 2015, ymin = -Inf, ymax = Inf, alpha = .2) +
  geom_point(data = ts_ex[ts_ex$shock_0.25==1,], col = "yellow") +
  ggtitle("Smooth with span of 0.25") +
  geom_hline(aes(yintercept = 0), linetype = "dashed")+
   theme(plot.title = element_text(size = 10)) +
  geom_smooth(span = 0.25, col = "yellow", level = 0.95)

# signal-to-noise ratio
plot_ts_10th <- ggplot(ts_ex, aes(x = years, y = percent_diff)) + geom_point() + geom_line() +
  theme_bw() + xlab("Years") + ylab("% Difference") +
  annotate("rect", xmin = 1983, xmax = 2015, ymin = -Inf, ymax = Inf, alpha = .2) +
  geom_segment(aes(x = 2070, y = avg, xend = 2099, yend = avg), colour = "forestgreen") +
  ggtitle("End of century average beyond 10th percentile of historical period") +
  theme(legend.position = "none") +
  geom_hline(aes(yintercept = 0), linetype = "dashed") +
  annotate("rect", xmin = -Inf, xmax = Inf, ymin = quantile(ts_ex$percent_diff[1:33], probs = .10), ymax = quantile(ts_ex$percent_diff[1:33], probs = .90), fill = "lightgreen", alpha = .5) +
  geom_segment(aes(x = 2085, xend = 2085, y = quantile(ts_ex$percent_diff[1:33], probs = .1),
                   yend = avg), arrow = arrow(length = unit(0.5, "cm")),
               colour = "forestgreen", size = 0.5)+
   theme(plot.title = element_text(size = 10))

# summary panel plot
grid.arrange(plot_ts_avg, plot_ts_10th,
             plot_shock0.25, plot_shock0.5, plot_shock0.75, 
             nrow = 2)
```

## SSP effect on maize crops in the United States

```{r, warning=FALSE, message=FALSE, fig.width=10}

y <- ts %>% 
    filter(years<2100,
           output_variable=="maize",
           regions == "United States") %>% 
    data.frame() %>% 
    group_by(sector, spatial_scale, regions, eco_model, climate_model, experiment_climate, output_variable) %>% 
    summarize(percentile_90 = as.vector(quantile(percent_diff[1:33], 0.90, na.rm=TRUE)),
              percentile_10 = as.vector(quantile(percent_diff[1:33], 0.10, na.rm=TRUE)),
              percent_diff = mean(percent_diff[88:117], na.rm=T)) %>% 
    mutate(change_5 = ifelse(percent_diff>5 | percent_diff<(-5), 1, 0),
           change_10 = ifelse(percent_diff>10 | percent_diff<(-10), 1, 0),
           change_25 = ifelse(percent_diff>25 | percent_diff<(-25), 1, 0),
           change_pct = ifelse(percent_diff<percentile_10 | percent_diff>percentile_90, 1, 0)) %>% 
    group_by(sector, spatial_scale, regions, climate_model, experiment_climate, output_variable) %>% 
    summarize(change_5 = round(mean(change_5, na.rm=T),4),
              change_10 = round(mean(change_10, na.rm=T),4),
              change_25 = round(mean(change_25, na.rm=T),4),
              change_pct = round(mean(change_pct, na.rm=T),4)) %>% 
    mutate(climates = paste(climate_model, experiment_climate)) %>% 
    group_by(spatial_scale, regions, climates) %>% 
    summarize(at_least_one_change_5 = get_at_least_one(6, change_5),
              at_least_two_change_5 = get_at_least_two(6, change_5),
              at_least_three_change_5 = get_at_least_three(6, change_5),
              at_least_one_change_10 = get_at_least_one(6, change_10),
              at_least_two_change_10 = get_at_least_two(6, change_10),
              at_least_three_change_10 = get_at_least_three(6, change_10),
              at_least_one_change_25 = get_at_least_one(6, change_25),
              at_least_two_change_25 = get_at_least_two(6, change_25),
              at_least_three_change_25 = get_at_least_three(6, change_25),
              at_least_one_change_pct = get_at_least_one(6, change_pct),
              at_least_two_change_pct = get_at_least_two(6, change_pct),
              at_least_three_change_pct = get_at_least_three(6, change_pct)) %>% 
    pivot_longer(4:15, names_to = "type", values_to = "probas") %>% 
  mutate(n_res = ifelse(str_detect(type, "one")==TRUE, "1", NA),
         n_res = ifelse(str_detect(type, "two")==TRUE, "2", n_res),
         n_res = ifelse(str_detect(type, "three")==TRUE, "3", n_res),
         threshold = ifelse(str_detect(type, "5")==TRUE, "5", NA),
         threshold = ifelse(str_detect(type, "10")==TRUE, "10", threshold),
         threshold = ifelse(str_detect(type, "25")==TRUE, "25", threshold),
         threshold = ifelse(str_detect(type, "pct")==TRUE, "pct", threshold))
```

## How often are 10th percentile limits reached per sector?

To understand the effect of using the 10th percentile of the historic period, we look at the 10th percentile and the % difference by the end of the century. This plot is excluding outliers when the % difference is higher than 200 etc.

```{r percentiles, warning=FALSE, message=FALSE, fig.width=10}

percentiles <- ts %>% 
  group_by(id, sector, eco_model, climate_model, experiment_climate, 
           output_variable, regions) %>% 
  summarize(avg = mean(percent_diff[88:117], na.rm=T),
            pct = ifelse(avg<0, quantile(percent_diff[1:33], probs = .05, na.rm=T), NA),
            pct = ifelse(avg>0, quantile(percent_diff[1:33], probs = .95, na.rm=T), pct)) %>% 
  mutate(climates = paste(climate_model, experiment_climate, sep="-")) %>% 
  filter(!output_variable %in% c("tcb", "tws", "ptotww"))

ggplot(percentiles, aes(x = pct, y = avg, color = climates)) + geom_point(alpha = .5) +
  facet_wrap(~output_variable, scales = "free") +
  scale_colour_manual(values = c("plum","purple4","lightskyblue","blue")) +
  theme_bw() +
  xlab("10th percentile threshold") + ylab("% difference") +
  geom_hline(aes(yintercept = 0), linetype = "dashed") +
  geom_vline(aes(xintercept = 0), linetype = "dashed") +
  geom_abline(aes(intercept = 0, slope = 1)) +
  xlim(-100,200) + ylim(-100,200)
```

## Cross-sector aggregated metrics

To quantify the cross-sector change, we calculated:

1.  The probability that at least 1, 2, or 3 resources change (sub-panels on the plot) beyond the 5, 10, or 25% fixed threshold per country and climate conditions (coded as 5, 10, and 25 on the x-axis)
2.  The probability that at least 1, 2, or 3 resources change beyond the 10th percentile of the historical period per country and climate conditions (coded as pct on the x-axis)
3.  The probability that there are at least one year from 2070-2099 during which at least 1, 2, or 3 resource shocks per country and climate conditions (coded as shock on the x-axis)

```{r cross-sector probs, warning=FALSE, message=FALSE, fig.width=10, fig.height = 3}
y <- ts %>% 
    filter(years<2100,
           !output_variable %in% c("ptotww","tcb","tws")) %>% 
    data.frame() %>% 
    group_by(sector, spatial_scale, regions, eco_model, climate_model, experiment_climate, output_variable) %>% 
    summarize(percentile_90 = as.vector(quantile(percent_diff[1:33], 0.90, na.rm=TRUE)),
              percentile_10 = as.vector(quantile(percent_diff[1:33], 0.10, na.rm=TRUE)),
              percent_diff = mean(percent_diff[88:117], na.rm=T)) %>% 
    mutate(change_5 = ifelse(percent_diff>5 | percent_diff<(-5), 1, 0),
           change_10 = ifelse(percent_diff>10 | percent_diff<(-10), 1, 0),
           change_25 = ifelse(percent_diff>25 | percent_diff<(-25), 1, 0),
           change_pct = ifelse(percent_diff<percentile_10 | percent_diff>percentile_90, 1, 0)) %>% 
    group_by(sector, spatial_scale, regions, climate_model, experiment_climate, output_variable) %>% 
    summarize(change_5 = round(mean(change_5, na.rm=T),4),
              change_10 = round(mean(change_10, na.rm=T),4),
              change_25 = round(mean(change_25, na.rm=T),4),
              change_pct = round(mean(change_pct, na.rm=T),4)) %>% 
    mutate(climates = paste(climate_model, experiment_climate)) %>% 
    group_by(spatial_scale, regions, climates) %>% 
    summarize(at_least_one_change_5 = get_at_least_one(6, change_5),
              at_least_two_change_5 = get_at_least_two(6, change_5),
              at_least_three_change_5 = get_at_least_three(6, change_5),
              at_least_one_change_10 = get_at_least_one(6, change_10),
              at_least_two_change_10 = get_at_least_two(6, change_10),
              at_least_three_change_10 = get_at_least_three(6, change_10),
              at_least_one_change_25 = get_at_least_one(6, change_25),
              at_least_two_change_25 = get_at_least_two(6, change_25),
              at_least_three_change_25 = get_at_least_three(6, change_25),
              at_least_one_change_pct = get_at_least_one(6, change_pct),
              at_least_two_change_pct = get_at_least_two(6, change_pct),
              at_least_three_change_pct = get_at_least_three(6, change_pct)) %>% 
    pivot_longer(4:15, names_to = "type", values_to = "probas") %>% 
  mutate(n_res = ifelse(str_detect(type, "one")==TRUE, "1", NA),
         n_res = ifelse(str_detect(type, "two")==TRUE, "2", n_res),
         n_res = ifelse(str_detect(type, "three")==TRUE, "3", n_res),
         threshold = ifelse(str_detect(type, "5")==TRUE, "5", NA),
         threshold = ifelse(str_detect(type, "10")==TRUE, "10", threshold),
         threshold = ifelse(str_detect(type, "25")==TRUE, "25", threshold),
         threshold = ifelse(str_detect(type, "pct")==TRUE, "pct", threshold))
  
s <- read_csv("/Users/auroremaureaud/Library/CloudStorage/Dropbox/Minerva 2021 Climate Politics/Aurore/Resource synthesis/future_resource_change/data/short-term_change/countries_shocks_1985-2015_probas_spans.csv") %>% 
  filter(spans == "shock_0.75") %>% 
  pivot_longer(4:6, names_to = "type", values_to = "probas") %>% 
  mutate(spatial_scale = "regions",
         n_res = ifelse(str_detect(type, "one")==TRUE, "1", NA),
         n_res = ifelse(str_detect(type, "two")==TRUE, "2", n_res),
         n_res = ifelse(str_detect(type, "three")==TRUE, "3", n_res),
         threshold = "shock") %>% 
  dplyr::select(spatial_scale, regions, climates, type, probas, n_res, threshold)

yy <- rbind(y, s)

ggplot(yy, aes(x = threshold, y = probas, fill = climates)) + geom_boxplot() +
  facet_wrap(~factor(n_res)) + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = -45, hjust=0)) +
  scale_fill_manual(values = c("plum","purple4","lightskyblue","blue")) +
  scale_x_discrete(limits = c("5","10","25","pct","shock")) +
  xlab("Metric") + ylab("Country probabilities")

```

The plots below detail the correlation and relationships between the probability of at least two shock, the probability of at least 2 resources changing beyond 25% and the probability of at least 2 resources changing beyond the 10th or 90th percentile by the end of century by climate model and experiment. The Blue line is a loess smoother.

```{r relationships, warning=FALSE, message=FALSE, fig.width=7, fig.height = 3}

zz <- yy %>% 
  filter(type %in% c("at_least_two_change_25","at_least_two_change_pct","at_least_two_shocks")) %>% 
  dplyr::select(-threshold) %>% 
  pivot_wider(values_from = probas, names_from = type)

zz %>% 
  ggplot(aes(x = at_least_two_change_25, y = at_least_two_change_pct)) +
  geom_point() + theme_bw() +
  facet_wrap(~climates) +
  geom_smooth(stats="loess") +
  geom_abline(aes(intercept=0, slope=1), linetype="dashed")

zz %>% 
  ggplot(aes(x = at_least_two_shocks, y = at_least_two_change_pct)) +
  geom_point() + theme_bw() +
  facet_wrap(~climates) +
  geom_smooth(stats="loess") +
  geom_abline(aes(intercept=0, slope=1), linetype="dashed")

zz %>% 
  ggplot(aes(x = at_least_two_shocks, y = at_least_two_change_25)) +
  geom_point() + theme_bw() +
  facet_wrap(~climates) +
  geom_smooth(stats="loess") +
  geom_abline(aes(intercept=0, slope=1), linetype="dashed")
```
