---
title: "resource shock detection"
date: "`r format(Sys.time(), '%B, %Y')`"
fontsize: 20pt
output:
  pdf_document:
    toc: false
    toc_depth: '3'
  html_document:
    toc: false
    toc_depth: '3'
    df_print: paged
  word_document:
    toc: false
    toc_depth: '3'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)

# load libraries
library(raster)
library(ncdf4)
library(here)
library(tidyverse)
library(ggplot2)
library(sf)
sf_use_s2(FALSE)
library(gridExtra)
library(pracma)
library(zoo)
```

Input on this methods from: Malin Pinsky, Thomas Frölicher

Coding: Aurore Maureaud

## Data

Example data picked for this test: agriculture sector, CROVER crop model, GFDL-ESM4 climate model, SSP585 climate experiment, and 4 agricultural crops: maize, wheat, soybean, rice.

For each method used, the residuals of the method are computed and z-scored, and shocks are identified as crop yields exceeding 2SD from the mean (below -2 and above 2 or the z-scored residuals).

```{r load data, echo=FALSE, message=FALSE, warning=FALSE}
dat <- read_csv("/Users/auroremaureaud/Library/CloudStorage/Dropbox/Minerva 2021 Climate Politics/Aurore/Resource synthesis/future_resource_change/data/long-term_change/global_time_series_2010-2019_2090-2099.csv") %>% 
  filter(experiment_climate == "ssp585",
         output_variable != "ptotww",
         climate_model == "gfdl-esm4")

sub_dat <- dat %>% 
  filter(eco_model == "crover")

```

## Detecting annual resource shocks

### Detrending with linear models

This method is most straightforward used in [Fredston et al., 2023](https://www.nature.com/articles/s41586-023-06449-y) but does not work well for non-linear trends (which is most cases in long-term data from ISIMIP). In the example shown below, on the maize crop data seems to be properly detrended by the linear model. This is also evident from the R-square of the model (95% for maize, and \<50% for other crops).

```{r lm, echo=FALSE, message=FALSE, warning=FALSE}

crops <- sort(unique(sub_dat$output_variable))

for(c in 1:length(crops)){
  # subset data
  xx <- sub_dat %>% 
    filter(output_variable==crops[c])
  
  # linear model
  lmd <- lm(xx$percent_diff ~ xx$years)
  
  # residuals from linear model
  xx$res <- (resid(lmd)-mean(resid(lmd)))/sd(resid(lmd))

  # plot model residuals and shocks
  assign(paste0("plot_",crops[c]),
         ggplot(xx, aes(x = years, y = res)) + geom_line() + 
  geom_point(data = xx[(xx$res<(-2) | xx$res>2),], aes(x = years, y = res), col = "red") +
  geom_text(data = xx[(xx$res<(-2) | xx$res>2),], aes(label = years), col = "red",
  vjust = -0.1, hjust = -0.1, size = 2) +
  geom_hline(yintercept = -2, linetype = "dashed", color = "grey50") +
  geom_hline(yintercept = 2, linetype = "dashed", color = "grey50") +
  theme_bw() +
  ggtitle(paste0(xx$eco_model[1], " ", xx$output_variable[1], "  R2=", round(summary(lmd)$adj.r.squared, 2))) +
  xlab("Years") + ylab("z-score of residuals") +
  theme(text = element_text(size = 10)))
  
}

grid.arrange(plot_maize, plot_wheat, plot_soybean, plot_rice,nrow = 2)
```

### 31-year moving average

A 31-year average works well but it might not apply to the entire time-series of the dataset. The rolling window of 31 is chosen as it eliminates most of the interannual (e.g., ENSO, NAO, etc) and decadal variability (PDO) in the climate/ocean system (it is also often used in IPCC assessments).

The rollmean() function from the \`zoo\` R package is used, where the k rolling window parameter needs to be chosen.

```{r movingaverage, echo=FALSE, message=FALSE, warning=FALSE}

crops <- sort(unique(sub_dat$output_variable))

for(c in 1:length(crops)){
  # subset data
  xx <- sub_dat %>% 
    filter(output_variable==crops[c])
  
  # linear model
  running <- rollmean(xx$percent_diff, k = 31)

  # residuals from rolling mean
  res <- xx$percent_diff[16:102]-running
  xx$res <- NA
  xx$res[16:102] <- (res-mean(res))/sd(res)

  # plot model residuals and shocks
  assign(paste0("plot_",crops[c]),
         ggplot(xx, aes(x = years, y = res)) + geom_line() + 
  geom_point(data = xx[(xx$res<(-2) | xx$res>2),], aes(x = years, y = res), col = "red") +
  geom_text(data = xx[(xx$res<(-2) | xx$res>2),], aes(label = years), col = "red",
  vjust = -0.1, hjust = -0.1, size = 2) +
  geom_hline(yintercept = -2, linetype = "dashed", color = "grey50") +
  geom_hline(yintercept = 2, linetype = "dashed", color = "grey50") +
  theme_bw() +
  ggtitle(paste0(xx$eco_model[1], " ", xx$output_variable[1])) +
  xlab("Years") + ylab("z-score of residuals") +
  theme(text = element_text(size = 10)))
  
}

grid.arrange(plot_maize, plot_wheat, plot_soybean, plot_rice,nrow = 2)
```

An extrapolation of the time-series could be considered, as done in a previous [publication](https://doi.org/10.5194/egusphere-2024-488): "The past anthropogenic warming is here estimated by applying a 31-year running mean with a linear extrapolation for the last 15 years, assuming constant warming rate based on the last 31 years."

### Savitzky-Golay

A Savitzky-Golay low-pass filter (e.g., as in [this paper](https://doi.org/10.1016/j.sigpro.2005.02.002)) can detrend the data to isolate shocks.

-   See brief description of low-pass filters [here](https://se.mathworks.com/discovery/low-pass-filter.html).

<!-- -->

-   See brief description of Savitzky-Golay filters [here](https://se.mathworks.com/help/signal/ref/sgolayfilt.html).

The savgol() function from the \`pracma\` R package is used, where the derivative order parameter *dorder* is set to 0 for smoothing, the filter order parameter *forder* is set to 4 for a quartic filter, and the filter length parameter *fl* is set at 11.

```{r savitskygolay, echo=FALSE, message=FALSE, warning=FALSE}
crops <- sort(unique(sub_dat$output_variable))

for(c in 1:length(crops)){
  # subset data
  xx <- sub_dat %>% 
    filter(output_variable==crops[c])
  
  # savistky-golay with 
    low_pass <- savgol(xx$percent_diff, 11, 
                       dorder = 0, forder = 4)
      
    # plot(xx$percent_diff ~ xx$years, col = "grey")
    # lines(low_pass ~ xx$years)
    
    low_pass_res <- xx$percent_diff-low_pass
    xx$res <- (low_pass_res-mean(low_pass_res))/(sd(low_pass_res))

  # plot model residuals and shocks
  assign(paste0("plot_",crops[c]),
         ggplot(xx, aes(x = years, y = res)) + geom_line() + 
  geom_point(data = xx[(xx$res<(-2) | xx$res>2),], aes(x = years, y = res), col = "red") +
  geom_text(data = xx[(xx$res<(-2) | xx$res>2),], aes(label = years), col = "red",
  vjust = -0.1, hjust = -0.1, size = 2) +
  geom_hline(yintercept = -2, linetype = "dashed", color = "grey50") +
  geom_hline(yintercept = 2, linetype = "dashed", color = "grey50") +
  theme_bw() +
  ggtitle(paste0(xx$eco_model[1], " ", xx$output_variable[1])) +
  xlab("Years") + ylab("z-score of residuals") +
  theme(text = element_text(size = 10)))
  
}

grid.arrange(plot_maize, plot_wheat, plot_soybean, plot_rice,nrow = 2)
```

#### Sensitivity to filter length

Filter order is set at 11.

```{r filterspan, echo=FALSE, warning=FALSE, message=FALSE}
xx <- sub_dat %>% 
  filter(output_variable=="maize")

spans <- c(7, 9, 11, 15)
for(s in 1:length(spans)){
  
  # savistky-golay with 
    low_pass <- savgol(xx$percent_diff, fl = spans[s], 
                       dorder = 0, forder = 4)
    
    low_pass_res <- xx$percent_diff-low_pass
    xx$res <- (low_pass_res-mean(low_pass_res))/(sd(low_pass_res))
  
    # plot model residuals and shocks
  assign(paste0("plot_",spans[s]),
         ggplot(xx, aes(x = years, y = res)) + geom_line(col = "grey80") + 
  geom_point(data = xx[(xx$res<(-2) | xx$res>2),], aes(x = years, y = res), col = "grey20") +
  geom_text(data = xx[(xx$res<(-2) | xx$res>2),], aes(label = years), col = "grey20",
  vjust = -0.1, hjust = -0.1, size = 2) +
  geom_hline(yintercept = -2, linetype = "dashed", color = "grey50") +
  geom_hline(yintercept = 2, linetype = "dashed", color = "grey50") +
  theme_bw() +
  ggtitle(paste0(xx$eco_model[1], " ", xx$output_variable[1], "  span=",spans[s])) +
  xlab("Years") + ylab("z-score of residuals") +
  theme(text = element_text(size = 10)))
}

grid.arrange(plot_7, plot_9, plot_11, plot_15,nrow = 2)
```

#### Sensitivity to the filter order and length

Filter is either quadratic (2), or quartic (4), and filter length is either set a 7 or 11.

```{r filterorder, echo=FALSE, warning=FALSE, message=FALSE}
xx <- sub_dat %>% 
  filter(output_variable=="maize")

orders <- c(2,4)
spans <- c(7,11)
for(o in 1:length(orders)){
  for(s in 1:length(spans)){
  
  # savistky-golay with 
    low_pass <- savgol(xx$percent_diff, fl = spans[s], 
                       dorder = 0, forder = orders[o])
    
    low_pass_res <- xx$percent_diff-low_pass
    xx$res <- (low_pass_res-mean(low_pass_res))/(sd(low_pass_res))
  
    # plot model residuals and shocks
  assign(paste0("plot_",orders[o],"_",spans[s]),
         ggplot(xx, aes(x = years, y = res)) + geom_line(col = "grey50") + 
  geom_point(data = xx[(xx$res<(-2) | xx$res>2),], aes(x = years, y = res), col = "grey20") +
  geom_text(data = xx[(xx$res<(-2) | xx$res>2),], aes(label = years), col = "grey20",
  vjust = -0.1, hjust = -0.1, size = 2) +
  geom_hline(yintercept = -2, linetype = "dashed", color = "grey50") +
  geom_hline(yintercept = 2, linetype = "dashed", color = "grey50") +
  theme_bw() +
  ggtitle(paste0(xx$eco_model[1], " ", xx$output_variable[1], "  order=",orders[o]," span=",spans[s])) +
  xlab("Years") + ylab("z-score of residuals") +
  theme(text = element_text(size = 10)))
  }
}

grid.arrange(plot_2_7, plot_2_11, plot_4_7, plot_4_11,ncol =2)
```

### Smoother across crops

The loess() function from the \`stats\` R package is used, where the parameter controlling the degree of smoothing span is set at 1, and the degree of the polynomial is 2.

```{r smooth, echo=FALSE, message=FALSE, warning=FALSE}

crops <- sort(unique(sub_dat$output_variable))

for(c in 1:length(crops)){
  # subset data
  xx <- sub_dat %>% 
    filter(output_variable==crops[c])
  
  # loess with span = 1
    model_smooth <- loess(xx$percent_diff ~ xx$years, span =1)
  xx$res <- (model_smooth$residuals - mean(model_smooth$residuals))/sd(model_smooth$residuals)
  
  # plot model residuals and shocks
  assign(paste0("plot_",crops[c]),
         ggplot(xx, aes(x = years, y = res)) + geom_line() + 
  geom_point(data = xx[(xx$res<(-2) | xx$res>2),], aes(x = years, y = res), col = "red") +
  geom_text(data = xx[(xx$res<(-2) | xx$res>2),], aes(label = years), col = "red",
  vjust = -0.1, hjust = -0.1, size = 2) +
  geom_hline(yintercept = -2, linetype = "dashed", color = "grey50") +
  geom_hline(yintercept = 2, linetype = "dashed", color = "grey50") +
  theme_bw() +
  ggtitle(paste0(xx$eco_model[1], " ", xx$output_variable[1])) +
  xlab("Years") + ylab("z-score of residuals") +
  theme(text = element_text(size = 10)))
  
}

grid.arrange(plot_maize, plot_wheat, plot_soybean, plot_rice,nrow = 2)
```

#### Sensitivity to smoother span

Testing the sensitivity of the shock detection to the smoother span, we can see that for the maize model example, we detect 5, 6, or 7 shocks for a span from 0.5 to 1.5.

```{r smoothspan, echo=FALSE, message=FALSE, warning=FALSE}
xx <- sub_dat %>% 
  filter(output_variable=="maize")

spans <- c(0.5, 0.75, 1, 1.5)
for(s in 1:length(spans)){
 
  # loess with span s
    model_smooth <- loess(xx$percent_diff ~ xx$years, span =spans[s])
  xx$res <- (model_smooth$residuals - mean(model_smooth$residuals))/sd(model_smooth$residuals)
  
    # plot model residuals and shocks
  assign(paste0("plot_",spans[s]),
         ggplot(xx, aes(x = years, y = res)) + geom_line(col = "grey80") + 
  geom_point(data = xx[(xx$res<(-2) | xx$res>2),], aes(x = years, y = res), col = "grey20") +
  geom_text(data = xx[(xx$res<(-2) | xx$res>2),], aes(label = years), col = "grey20",
  vjust = -0.1, hjust = -0.1, size = 2) +
  geom_hline(yintercept = -2, linetype = "dashed", color = "grey50") +
  geom_hline(yintercept = 2, linetype = "dashed", color = "grey50") +
  theme_bw() +
  ggtitle(paste0(xx$eco_model[1], " ", xx$output_variable[1], "  span=",spans[s])) +
  xlab("Years") + ylab("z-score of residuals") +
  theme(text = element_text(size = 10)))
}

grid.arrange(plot_0.5, plot_0.75, plot_1, plot_1.5,nrow = 2)
```

#### Sensitivity to smoother span and degree

Sensitivity to the smoother span and polynomial degree.

```{r smoothspandegree, echo=FALSE, message=FALSE, warning=FALSE}

xx <- sub_dat %>% 
  filter(output_variable=="maize")

spans <- c(0.75, 1.5)
degrees <- c(1,2)
for(s in 1:length(spans)){
  for(d in 1:length(degrees)){
 
  # loess with span s
    model_smooth <- loess(xx$percent_diff ~ xx$years, span =spans[s], degree = degrees[d])
  xx$res <- (model_smooth$residuals - mean(model_smooth$residuals))/sd(model_smooth$residuals)
  
    # plot model residuals and shocks
  assign(paste0("plot_",spans[s],"_",degrees[d]),
         ggplot(xx, aes(x = years, y = res)) + geom_line(col = "grey50") + 
  geom_point(data = xx[(xx$res<(-2) | xx$res>2),], aes(x = years, y = res), col = "grey20") +
  geom_text(data = xx[(xx$res<(-2) | xx$res>2),], aes(label = years), col = "grey20",
  vjust = -0.1, hjust = -0.1, size = 2) +
  geom_hline(yintercept = -2, linetype = "dashed", color = "grey50") +
  geom_hline(yintercept = 2, linetype = "dashed", color = "grey50") +
  theme_bw() +
  ggtitle(paste0(xx$eco_model[1], " ", xx$output_variable[1], "  span=",spans[s]," degree=",degrees[d])) +
  xlab("Years") + ylab("z-score of residuals") +
  theme(text = element_text(size = 10)))
  }
}

grid.arrange(plot_0.75_1, plot_0.75_2, plot_1.5_1, plot_1.5_2,nrow = 2)
```
