---
title: "Baseline shock probabilities"
date: "`r format(Sys.time(), '%B, %Y')`"
fontsize: 20pt
output:
  html_document:
    toc: false
    toc_depth: '3'
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

rm(list = ls())

# load libraries
library(here)
library(tidyverse)
library(ggplot2)
library(gridExtra)
library(sf)
library(colorspace)
library(ggpubr)

# load functions
source("/Users/auroremaureaud/Library/CloudStorage/Dropbox/Minerva 2021 Climate Politics/Aurore/Resource synthesis/future_resource_change/functions/get_probas.R")

# load future shock probabilities
f_shocks <- read_csv("/Users/auroremaureaud/Library/CloudStorage/Dropbox/Minerva 2021 Climate Politics/Aurore/Resource synthesis/future_resource_change/data/short-term_change/countries_shocks_1985-2015_probas_spans.csv") %>% 
  dplyr::select(regions, climates, spans, at_least_two_shocks) %>% 
  rename(pshocks_f = at_least_two_shocks)

# calculate past shock probabilities
xx <- read_csv("/Users/auroremaureaud/Library/CloudStorage/Dropbox/Minerva 2021 Climate Politics/Aurore/Resource synthesis/future_resource_change/data/short-term_change/countries_shocks_time_series_1985-2015_spans.csv")%>% 
  filter(!output_variable %in% c("tws","tcb","ptotww")) %>%
  mutate(climates = paste(climate_model, experiment_climate)) %>%
  group_by(regions, climates, years, output_variable) %>% 
  summarize(shock_0.75 = round(mean(shock_0.75, na.rm=T),4),
            shock_0.75_down = round(mean(shock_0.75_down, na.rm=T),4),
            shock_0.75_up = round(mean(shock_0.75_up, na.rm=T),4),
            shock_0.5 = round(mean(shock_0.5, na.rm=T),4),
            shock_0.5_down = round(mean(shock_0.5_down, na.rm=T),4),
            shock_0.5_up = round(mean(shock_0.5_up, na.rm=T),4),
            shock_0.25 = round(mean(shock_0.25, na.rm=T),4),
            shock_0.25_down = round(mean(shock_0.25_down, na.rm=T),4),
            shock_0.25_up = round(mean(shock_0.25_up, na.rm=T),4)) %>% 
  dplyr::select(regions, climates, years, output_variable, shock_0.75, shock_0.5, shock_0.25) %>% 
  pivot_longer(5:7, names_to = "spans", values_to = "shock") %>% 
  pivot_wider(names_from = output_variable, values_from = "shock") %>% 
  data.frame() 

spans <- c("0.25","0.5","0.75")
shock_proba_r_cs <- data.frame()
for(s in 1:length(spans)){
  
  span <- spans[s]
  yy <- xx %>% 
    filter(years>1984,
           years<2016,
           str_detect(spans, span)==TRUE)
  
  yy$at_least_one_shock <- yy$at_least_two_shocks <- yy$at_least_three_shocks <- NA
  yy$at_least_one_shock <- apply(yy[,5:10], 1, function(x) get_at_least_one(6,x))
  yy$at_least_two_shocks <- apply(yy[,5:10], 1, function(x) get_at_least_two(6,x))
  yy$at_least_three_shocks <- apply(yy[,5:10], 1, function(x) get_at_least_three(6,x))
  
  yy <- yy %>% 
    dplyr::select(regions, climates, years, spans, at_least_one_shock, at_least_two_shocks, at_least_three_shocks) %>% 
    group_by(regions, climates, spans) %>% 
    summarize(at_least_one_shock = get_at_least_one(30,at_least_one_shock),
              at_least_two_shocks = get_at_least_one(30, at_least_two_shocks),
              at_least_three_shocks = get_at_least_one(30, at_least_three_shocks),
              time_window = 30)
  
  if(s==1){shock_proba_r_cs <- yy
  } else {shock_proba_r_cs <- rbind(shock_proba_r_cs,yy)}
  rm(yy)
}

p_shocks <- shock_proba_r_cs %>% 
  dplyr::select(regions, climates, spans, at_least_two_shocks) %>% 
  rename(pshocks_p = at_least_two_shocks)

# join the past and future shock probabilities
shocks <- left_join(f_shocks, p_shocks, by = c("regions","climates","spans")) %>% 
  filter(spans == "shock_0.75",
         climates %in% c("gfdl-esm4 ssp585","ipsl-cm6a-lr ssp585")) %>% 
  group_by(regions) %>% 
  summarize (pshocks_f = median(pshocks_f, na.rm=T),
             pshocks_p = median(pshocks_p, na.rm=T))

rm(span, spans, s, xx)

# load regions shapefile
regions <- st_read("/Users/auroremaureaud/Library/CloudStorage/Dropbox/Minerva 2021 Climate Politics/Aurore/Resource synthesis/future_resource_change/data/EEZ_land_union_v3_202003/EEZ_Land_v3_202030.shp") %>% 
  filter(POL_TYPE != "Joint regime (EEZ)",
         is.na(SOVEREIGN2),
         SOVEREIGN1 != "Republic of Mauritius",
         UNION != "Antarctica") %>% 
  dplyr::select(UNION, SOVEREIGN1)

# join the shapefile and data
shocks_map <- left_join(regions, shocks, by = c("SOVEREIGN1" = "regions"))
```

Mapping out the probability of at least two shocks across 2070-2099 for the SSP 5.85 climate scenario:

```{r map future shock proba, echo=FALSE, warning=FALSE, message=FALSE, fig.width=10, fig.height = 5}
ggplot(shocks_map) + geom_sf(aes(fill = pshocks_f)) +
  scale_fill_continuous_sequential(palette = "PuBuGn", expand = c(0,0)) +
    theme_bw() +
    theme(legend.position = "bottom",
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        strip.background = element_rect(color = "white", fill = "white"),
        panel.spacing = unit(0.25,'lines'))
```

Mapping out the probability of at least two shocks across 1985-2015:

```{r map past shock proba, echo=FALSE, warning=FALSE, message=FALSE, fig.width=10, fig.height = 5}
ggplot(shocks_map) + geom_sf(aes(fill = pshocks_p)) +
  scale_fill_continuous_sequential(palette = "PuBuGn", expand = c(0,0)) +
    theme_bw() +
    theme(legend.position = "bottom",
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        strip.background = element_rect(color = "white", fill = "white"),
        panel.spacing = unit(0.25,'lines'))
```

Relationship between past and future shock probability under SSP 5.85:

```{r relationship past-future per country, echo=FALSE, warning=FALSE, message=FALSE}

ggplot(shocks, aes(x = pshocks_p, y = pshocks_f)) + geom_point() +
  stat_smooth(method = "lm", col = "black") +
  theme_bw() +
  xlab("Probability 2 shocks+ 1985-2015") +
  ylab("Probability 2 shocks+ 2070-2100") +
  geom_abline(intercept = 0, slope = 1, lwd=1, linetype = "dashed",
              col = "red")
```

*The dark line and shaded gray area show a linear model fit, and the dotted red line shows the 1-to-1 line.*

Mapping out the probability of at least two shocks across 2070-2100 for SSP 1.26:

```{r map ssp126, echo=FALSE, warning=FALSE, message=FALSE, fig.width=10, fig.height = 5}
p_shocks <- shock_proba_r_cs %>% 
  dplyr::select(regions, climates, spans, at_least_two_shocks) %>% 
  rename(pshocks_p = at_least_two_shocks)

# join the past and future shock probabilities
shocks <- left_join(f_shocks, p_shocks, by = c("regions","climates","spans")) %>% 
  filter(spans == "shock_0.75",
         climates %in% c("gfdl-esm4 ssp126","ipsl-cm6a-lr ssp126")) %>%
  group_by(regions) %>% 
  summarize (pshocks_f = median(pshocks_f, na.rm=T),
             pshocks_p = median(pshocks_p, na.rm=T))

ggplot(shocks_map) + geom_sf(aes(fill = pshocks_p)) +
  scale_fill_continuous_sequential(palette = "PuBuGn", expand = c(0,0)) +
    theme_bw() +
    theme(legend.position = "bottom",
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        strip.background = element_rect(color = "white", fill = "white"),
        panel.spacing = unit(0.25,'lines'))
```

Relationship between past and future shock probabilities under SSP 1.26:
