---
title: "Compare p(X>=2) and p(X>=50%)"
date: "`r format(Sys.time(), '%B, %Y')`"
fontsize: 20pt
output:
  html_document:
    toc: false
    toc_depth: '3'
    df_print: paged
---

Metrics compared for shocks:

1. number of resources per country
2. probability of at least two shocks p(X>=2)
3. probability of at least 50% of shocks p(X>=50%). When there are 2 resources, p(X>=1) is applied, when there are 3-4 resources, p(X>=2) is applied and when there are 5-6 resources p(X>=3) is applied.
4. probability of at least 50% of shocks p(X>=50%)bis. When there are 2 resources, p(X>=1) is applied, when there are 3 resources mean(p(X>=1),p(X>=2) is applied, when there are 4 resources p(X>=2) is applied, when there are 5 resources mean(p(X>=2),p(X>=3)) is applied, and when there are 6 resources p(X>=3) is applied.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

rm(list = ls())

# load libraries
library(here)
library(tidyverse)
library(ggplot2)
library(gridExtra)
library(sf)
library(colorspace)
library(ggpubr)
library(GGally)

# load p(X>=2)
shocks_two <- read_csv("/Users/auroremaureaud/Library/CloudStorage/Dropbox/Minerva 2021 Climate Politics/Aurore/Resource synthesis/future_resource_change/data/short-term_change/countries_shocks_1985-2015_probas_spans.csv") %>% 
  filter(spans == "shock_0.75") %>% 
  dplyr::select(regions, climates, at_least_two_shocks)

# load p(X>=50%)
shocks_half <- read_csv("/Users/auroremaureaud/Library/CloudStorage/Dropbox/Minerva 2021 Climate Politics/Aurore/Resource synthesis/future_resource_change/data/short-term_change/countries_shocks_1985-2015_probas_spans_half.csv") %>% 
  filter(spans == "shock_0.75") %>% 
  dplyr::select(regions, climates, at_least_half_shocks, at_least_half_bis_shocks)

# join the past and future shock probabilities
shocks <- left_join(shocks_two, shocks_half, by = c("regions","climates")) %>% 
  filter(climates %in% c("gfdl-esm4 ssp585","ipsl-cm6a-lr ssp585")) %>% 
  pivot_longer(3:5, names_to = "type", values_to = "proba") %>% 
  group_by(regions, type) %>% 
  summarize(proba = median(proba, na.rm=T)) %>% 
  pivot_wider(values_from = proba, names_from = type)

# number of resources
pct_diff_ts_r_cs_ag <- read_csv("/Users/auroremaureaud/Library/CloudStorage/Dropbox/Minerva 2021 Climate Politics/Aurore/Resource synthesis/future_resource_change/data/data_processing/countries_time_series_1985-2015_2070-2099_agriculture.csv")
pct_diff_ts_r_cs_fi <- read_csv("/Users/auroremaureaud/Library/CloudStorage/Dropbox/Minerva 2021 Climate Politics/Aurore/Resource synthesis/future_resource_change/data/data_processing/countries_time-series_1985-2015_2070-2099_marine-fishery_global.csv")
pct_diff_ts_r_cs_wa <- read_csv("/Users/auroremaureaud/Library/CloudStorage/Dropbox/Minerva 2021 Climate Politics/Aurore/Resource synthesis/future_resource_change/data/data_processing/countries_time_series_1985-2015_2070-2099_water_global.csv")
pct_diff_ts_r_cs <- rbind(pct_diff_ts_r_cs_ag, pct_diff_ts_r_cs_fi, pct_diff_ts_r_cs_wa)

nres_r <- data.frame(pct_diff_ts_r_cs) %>% 
  filter(!is.na(percent_diff),
         output_variable %in% c("rice", "wheat", "soybean", "maize", "tcblog10", "qtot")) %>% 
  group_by(regions) %>% 
  summarize(n_res = length(unique(output_variable))/6)

shocks <- left_join(shocks, nres_r, by = "regions")

# load regions shapefile
regions <- st_read("/Users/auroremaureaud/Library/CloudStorage/Dropbox/Minerva 2021 Climate Politics/Aurore/Resource synthesis/future_resource_change/data/EEZ_land_union_v3_202003/EEZ_Land_v3_202030.shp") %>% 
  filter(POL_TYPE != "Joint regime (EEZ)",
         is.na(SOVEREIGN2),
         SOVEREIGN1 != "Republic of Mauritius",
         UNION != "Antarctica") %>% 
  dplyr::select(UNION, SOVEREIGN1)

# join the shapefile and data
shocks_map <- left_join(regions, shocks, by = c("SOVEREIGN1" = "regions"))
```

Mapping out the p(X>=2), p(X>=50%), and p(X>=50%)bis for the SSP 5.85 climate scenario for shocks:

```{r map shocks, echo=FALSE, warning=FALSE, message=FALSE, fig.width=10, fig.height=6}
shocks_map %>% 
  rename(`p(X>=2)` = at_least_two_shocks,
         `p(X>=50%)` = at_least_half_shocks,
         `p(X>=50%)bis` = at_least_half_bis_shocks,
         `# resources` = n_res) %>% 
  pivot_longer(3:6, names_to = "type", values_to = "proba") %>% 
ggplot() + geom_sf(aes(fill = proba)) +
  scale_fill_continuous_sequential(palette = "PuBuGn", expand = c(0,0)) +
  facet_wrap(~type, nrow = 2) +
    theme_bw() +
    theme(legend.position = "bottom",
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        strip.background = element_rect(color = "white", fill = "white"),
        panel.spacing = unit(0.25,'lines')) +
  ggtitle("Shock metrics")
```

Correlation table of shock metrics:

```{r cor shocks, echo=FALSE, warning=FALSE, message=FALSE, fig.width=10, fig.height=6}

ggpairs(shocks, columns = c(2:5)) +
  theme_bw() + theme(text = element_text(size = 16))

```

Metrics compared for gradual changes:

1. number of resources per country
2. probability of at least two resources changing beyond 25% p(X>=2)
3. probability of at least 50% of resources changing beyond 25% p(X>=50%). When there are 2 resources, p(X>=1) is applied, when there are 3-4 resources, p(X>=2) is applied and when there are 5-6 resources p(X>=3) is applied.
4. probability of at least 50% of resources changing beyond 25% p(X>=50%)bis. When there are 2 resources, p(X>=1) is applied, when there are 3 resources mean(p(X>=1),p(X>=2) is applied, when there are 4 resources p(X>=2) is applied, when there are 5 resources mean(p(X>=2),p(X>=3)) is applied, and when there are 6 resources p(X>=3) is applied.

Mapping out the p(X>=2), p(X>=50%), and p(X>=50%)bis for the SSP 5.85 climate scenario for gradual changes: 

```{r map gradual, echo=FALSE, warning=FALSE, message=FALSE, fig.width=10, fig.height=6}

# load p(X>=2)
gradual_two <- read_csv("/Users/auroremaureaud/Library/CloudStorage/Dropbox/Minerva 2021 Climate Politics/Aurore/Resource synthesis/future_resource_change/data/long-term_change/countries_long-term_change_1985-2015_probas.csv") %>% 
  filter(time_window == 30,
         type == "at_least_two_change_25") %>% 
  dplyr::select(regions, climates, type, probas) %>% 
  pivot_wider(names_from = type, values_from = probas)

# load p(X>=50%)
gradual_half <- read_csv("/Users/auroremaureaud/Library/CloudStorage/Dropbox/Minerva 2021 Climate Politics/Aurore/Resource synthesis/future_resource_change/data/long-term_change/countries_long-term_change_1985-2015_probas_half.csv") %>% 
  filter(time_window == 30,
         type %in% c("at_least_half_change_25","at_least_half_bis_change_25")) %>%
  dplyr::select(regions, climates, type, probas) %>% 
  pivot_wider(names_from = type, values_from = probas)

# join the past and future shock probabilities
gradual <- left_join(gradual_two, gradual_half, by = c("regions","climates")) %>% 
  filter(climates %in% c("gfdl-esm4 ssp585","ipsl-cm6a-lr ssp585")) %>% 
  pivot_longer(3:5, names_to = "type", values_to = "proba") %>% 
  group_by(regions, type) %>% 
  summarize(proba = median(proba, na.rm=T)) %>% 
  pivot_wider(values_from = proba, names_from = type)

# n_res
gradual <- left_join(gradual, nres_r, by = "regions")

# join the shapefile and data
gradual_map <- left_join(regions, gradual, by = c("SOVEREIGN1" = "regions"))

# map
gradual_map %>% 
  rename(`p(X>=2)` = at_least_two_change_25,
         `p(X>=50%)` = at_least_half_change_25,
         `p(X>=50%)bis` = at_least_half_bis_change_25,
         `# resources` = n_res) %>% 
  pivot_longer(3:6, names_to = "type", values_to = "proba") %>% 
ggplot() + geom_sf(aes(fill = proba)) +
  scale_fill_continuous_sequential(palette = "PuBuGn", expand = c(0,0)) +
  facet_wrap(~type, nrow = 2) +
    theme_bw() +
    theme(legend.position = "bottom",
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        strip.background = element_rect(color = "white", fill = "white"),
        panel.spacing = unit(0.25,'lines')) +
  ggtitle("Gradual metrics")

```

Correlation table of gradual metrics:

```{r cor gradual, echo=FALSE, warning=FALSE, message=FALSE}

ggpairs(gradual, columns = c(2:5)) +
  theme_bw() + theme(text = element_text(size = 16))

```
